---
title: "Texas Under-18 Population Growth"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})

---

This script uses R's tidycensus library to download county population data to reveal trends in the under-18 population, by ethnicity, from 2010 to 2017. Texas was a leader in under-18 population growth during that time. We also look at growth projections from the Texas Demographic Center, predicting those growth trends to continue for the foreseeable future.


```{r}
library(tidyverse)
library(tidycensus)
options(tigris_use_cache = TRUE)
```

## Under-18 population growth in Texas and other states, 2010-2017

First, compare counties in the six biggest states to see how Texas counties stack up in under-18 population growth.

```{r}
big10 <- get_acs("county", variables="B09001_001", year = 2010, output="tidy", state=c("Texas","California","New York","Illinois","Pennsylvania","Florida"), geometry=FALSE)
big17 <- get_acs("county", variables="B09001_001", year = 2017, output="tidy", state=c("Texas","California","New York","Illinois","Pennsylvania","Florida"), geometry=FALSE)
big10 <- big10 %>% rename('e10' = 'estimate') %>% select(c(-variable, -moe))
big17 <- big17 %>% rename('e17' = 'estimate') %>% select(c(-variable, -moe))
big <- inner_join(big10, big17, by="NAME")
big <- mutate(big, diff = e17 - e10)
big <- mutate(big, pct = (e17 / e10) - 1)
big <- arrange(big, desc(diff))
big[1:20,]
```

Just counties in Texas.

```{r}
texas_under18 <- filter(big, str_detect(NAME, 'Texas')) %>% arrange(desc(diff))
texas_under18[1:20,]
```

And counties in all states.

```{r}
all10 <- get_acs("county", variables="B09001_001", year = 2010, output="tidy", geometry=FALSE)
all17 <- get_acs("county", variables="B09001_001", year = 2017, output="tidy", geometry=FALSE)
all10 <- all10 %>% rename('e10' = 'estimate') %>% select(c(-variable, -moe))
all17 <- all17 %>% rename('e17' = 'estimate') %>% select(c(-variable, -moe))
all <- inner_join(all10, all17, by="NAME")
all <- mutate(all, diff = e17 - e10)
all <- arrange(all, desc(diff))
all[1:20,]
```
## Mobility of the under-18 population

Next we look at mobility/migration from other states and other countries, using summaries of B07001_002 (ages 0-4) and 003 (ages 5-17), and variables 066 and 067 (moved from different state) and 082 and 083 (moved from abroad). The goal is to produce a table with each county and fields for total 0-17, # migrated from other state, and number migrated from another country. This can be adapted to compare all counties in all states simply by removing the "state" parameter from the get_acs() call.

```{r}
migr17 <- get_acs("county", variables = c("B07001_002", "B07001_003", "B07001_066", "B07001_082", "B07001_067", "B07001_083"), year = 2017, output = "tidy", state = 48, geometry = FALSE)
migr17 <- migr17 %>% select(-moe) %>% spread(variable, estimate)
migr17 <- migr17 %>% mutate(total_kids = B07001_002 + B07001_003)
migr17 <- migr17 %>% mutate(another_state = B07001_066 + B07001_067)
migr17 <- migr17 %>% mutate(another_country = B07001_082 + B07001_083)
migr17 <- migr17 %>% select(-starts_with('B07'))
migr17 <- migr17 %>% mutate(pct_state = another_state / total_kids)
migr17 <- migr17 %>% mutate(pct_country = another_country / total_kids)
migr17 <- migr17 %>% mutate(total_migrants = another_state + another_country)
migr17 <- migr17 %>% mutate(pct_migrants = total_migrants / total_kids)
migr17 <- migr17 %>% arrange(desc(total_migrants))
migr17[1:20,]
```
## Texas Demographic Center projections of under-18 population, 2020-2050

Next we map the [Texas Demographic Center](https://demographics.texas.gov/Data/TPEPP/Projections/)'s projected differences by county and ethnicity between 2020 and 2050.

First we get the map geometry from ACS. Then we read in an Excel sheet, prepared from the projections provided on the TDC web site, and merge with the county geography. Finally we cycle through a series of maps to show where growth is expected, by raw numbers and percentages.

```{r}
under18 <- get_acs(state = "TX", geography = "county", variables = "B09001_001", geometry = TRUE, year = 2017)
under18$NAME <- sub(", Texas", "", under18$NAME)
library(readxl)
X2018allcntyagegrp <- read_excel("tdc_projections.xlsx", sheet = "diffs and pcts only")
counties <- inner_join(X2018allcntyagegrp, under18, by = c("Row Labels" = "NAME"))

ggplot(data = counties, aes(fill = pct_total)) + geom_sf() + scale_fill_gradient2(high = "dark blue")
ggplot(data = counties, aes(fill = pct_white)) + geom_sf() + scale_fill_gradient2(high = "dark blue")
ggplot(data = counties, aes(fill = pct_black)) + geom_sf() + scale_fill_gradient2(high = "dark blue")
ggplot(data = counties, aes(fill = pct_hisp)) + geom_sf() + scale_fill_gradient2(high = "dark blue")
ggplot(data = counties, aes(fill = pct_asian)) + geom_sf() + scale_fill_gradient2(high = "dark blue")
ggplot(data = counties, aes(fill = pct_other)) + geom_sf() + scale_fill_gradient2(high = "dark blue")

ggplot(data = counties, aes(fill = diff_white)) + geom_sf() + scale_fill_gradient2(high = "dark green")
ggplot(data = counties, aes(fill = diff_black)) + geom_sf() + scale_fill_gradient2(high = "dark green")
ggplot(data = counties, aes(fill = diff_hisp)) + geom_sf() + scale_fill_gradient2(high = "dark green")
ggplot(data = counties, aes(fill = diff_asian)) + geom_sf() + scale_fill_gradient2(high = "dark green")
ggplot(data = counties, aes(fill = diff_other)) + geom_sf() + scale_fill_gradient2(high = "dark green")
```

